Openness and Computational Reproducibility in plant pathology: where do we stand and a way forward

### 2022-06-04 analysis/paper/paper.Rmd
Edit the results section to match new analyses

All authors agreed on the evaluation of the computational methods and data availability of all five articles that were scored by each author (Figure \@ref(fig:irCriteria)A, \@ref(fig:irCriteria)B).

The analysis detected that the journal title did have an effect on the reproducibility score but there was no effect of article class or impact factor.

@@

We were unable to detect any observable differences in the years that publications were submitted.
The year 2016 appears to be have a higher score than prior and following years, but the analysis was undecided on the effect of all years on the publications' openness and reproducibility.

Most journals in our discipline fail to mention the availability of the computational methods to support the article as published.
Only two journals, *Plant Pathology* and *Tropical Plant Pathology* have full instructions for authors regarding both data and code deposition in repositories for sharing with readers.
Three had no instructions at all, *Forest Pathology*, *Phytopathologia Mediterranea* and *Revista Mexicana de Fitopatología*, with the remaining 16 journals having varying levels of instruction for authors to deposit data but not code.
While the analysis did not detect an effect of the journals' instructions for sharing or depositing data and or code, the more detailed the instructions, the higher the article scored.

The majority of the articles, `r length(which(rrpp$art_class == "Fundamental"))`, were classed as fundamental, with the remaining `r length(which(rrpp$art_class != "Fundamental"))` classed as applied.

Software citations were also reasonably well reported in the literature with `r length(which(rrpp$software_cite == 2))` (`r round(length(which(rrpp$software_cite == 2))/nrow(rrpp), 2) * 100`%) scoring '2' where the main software was cited with version numbers and `r length(which(rrpp$software_cite == 3))` (`r round(length(which(rrpp$software_cite == 3))/nrow(rrpp), 2) * 100`%) scoring '3' where full citations including R packages or the SAS PROC used were cited where appropriate (Figure \@ref(fig:articleScores)C).

However, data availability was mostly not available with `r length(which(rrpp$data_avail == 0))` (`r round(length(which(rrpp$data_avail == 0))/nrow(rrpp), 2) * 100`%) scoring '0' where the data was not available or mentioned in the article.
However, more articles, `r length(which(rrpp$data_avail == 3))` (`r round(length(which(rrpp$data_avail == 3))/nrow(rrpp), 2) * 100`%) scored '3' than scored '1' or '2' combined, `r length(which(rrpp$data_avail == 2 | rrpp$data_avail == 1))` (`r round(length(which(rrpp$data_avail == 2 | rrpp$data_avail == 1))/nrow(rrpp), 2) * 100`%) with `r as.character(english(length(which(is.na(rrpp$data_avail)))))` articles not producing shareable data (Figure \@ref(fig:articleScores)D).

>>>>

However, data was mostly not available with `r length(which(rrpp$data_avail == 0))` (`r round(length(which(rrpp$data_avail == 0))/nrow(rrpp), 2) * 100`%) scoring '0' where the data was not available or mentioned in the article.
However, more articles, `r length(which(rrpp$data_avail == 3))` (`r round(length(which(rrpp$data_avail == 3))/nrow(rrpp), 2) * 100`%) scored '3' than scored '1' or '2' combined, `r length(which(rrpp$data_avail == 2 | rrpp$data_avail == 1))` (`r round(length(which(rrpp$data_avail == 2 | rrpp$data_avail == 1))/nrow(rrpp), 2) * 100`%) with `r as.character(english(length(which(is.na(rrpp$data_avail)))))` articles not producing shareable data (Figure \@ref(fig:articleScores)D).

## Interrater Agreement

All authors agreed on the evaluation of the data availability of all five articles that were scored by each author (Figure \@ref(fig:irCriteria)A) but one of the authors rated one article as `NA` or no computational methods used rather than zero as with the other four authors (Figure \@ref(fig:irCriteria)B), giving the interrater percent agreement a score of 100% with the `NA value dropped from the code availability.
Fleiss' Kappa was not able to be calculated for code or data availability due to 100% agreement of a score of '0', converted to '1' for the analysis.

## Effect of Journal Title

The analysis found that publications in Phytopathology were more likely to share their code but was unable to detect differences in any other journal titles.
Publications in Crop Protection and Plant Disease were less likely to share their data but no effects were detectable in any other journal title.
The analysis was unable to detect any effect of journal title on the software used, though articles in Plant Disease, Nematology and Tropical Plant Pathology tended to be more open and available.

## Effect of Impact Factor

There were no detectable effects of the five-year impact factor on code or data availability.
However, there was no effect of the five-year impact factor on the availability of the software that was used.

## Effect of Year

We were unable to detect any observable differences as a result of the year that an article was published for code, data or software availability.

## Effect of Access Policy

@@

## Effect of Reproducibility Instructions

Most journals in our discipline fail to mention the availability of the computational methods to support the article as published.
Only two journals, *Plant Pathology* and *Tropical Plant Pathology* have full instructions for authors regarding both data and code deposition in repositories for sharing with readers.
Three had no instructions at all, *Forest Pathology*, *Phytopathologia Mediterranea* and *Revista Mexicana de Fitopatología*, with the remaining 16 journals having varying levels of instruction for authors to deposit data but not code.
There were no detectable effects of the instructions on any of the three scores that were evaluated.

## Effect of Software Used

The availability of the software used had no effect on the availability of the code, i.e., the use of free, open-source software (FOSS), did not have the effect of making the code more available.
Conversely, the use of closed or restrictive software did not cause code to be less available. 
There was no detectable effect of the software's availability on the availability of data.
However, both of these scores tended to be higher when the software used was more available.
